Fast and exact visibility on digitized shapes and application to saliency-aware normal estimation

### 2025-04-24 DGMM2025/paper.tex
Update figure 10
Reduce intro

\usepackage{changes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{varwidth}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage[dvipsnames]{xcolor}
\pgfplotsset{compat=1.18}
\usetikzlibrary {arrows.meta,decorations.shapes}
\setauthormarkup{}
\definechangesauthor[name=jo, color=blue]{JO}
\definechangesauthor[name=rom, color=teal]{ROM}

@@

    Visibility is a fundamental concept in computational geometry and
    digital topology, with applications ranging from computer vision
    (identifying occlusions) to geometric modeling (feature detection, geodesics)
    and computer graphics (path tracing, shadowmap). Visibility within
    polygons in the plane has been extensively studied in the
    litterature~\cite{ghosh:2007-book} and is still studied in higher
    dimensional spaces~\cite{orourke:2017-book}. The general problem
    has a high complexity. For instance, the 3d visibility complex
    between $n$ spheres and occluding spheres has a complexity
    $O(n^4)$~\cite{durand:2002-tog}. Exact 3d visibility between
    polygons and occluding polygons is often casted into 5d Euclidean
    space derived from Plücker space~\cite{nirenstein:2002-ewr}, which
    nicely represents 3d lines and their respective positions. The
    computational cost remains high even with decomposition into
    coarse cells for speed up (several days of computation for a
    million triangles at the time).

@@

    We focus on solving efficiently this problem because it is crucial
    in computing geometric normals along digital surfaces that are
    aware of salient features. Indeed, 2d discrete tangent estimators
    based on tangency or discrete line segments have proven to be both
    sensitive to sharp corners while multigrid convergent along the
    boundary of digitized smooth shapes~\cite{feschet:1999-dgci,lachaud:2007-ivc,nguyen:2011-pr}.
    Many ad hoc methods have been proposed for tangent plane or normal
    estimation on 3d digital surfaces
    (e.g. \cite{fourey:2009-cg,charrier:2011-iwcia,Cuel:2014-dgci,Lachaud:2017-lnm,mareche:2024-ispr}).
    If two of these methods~\cite{Cuel:2014-dgci,Lachaud:2017-lnm}
    have been shown to be multigrid convergent on digitization of
    shapes with smooth boundary, their formulation intrinsically
    smoothes features, e.g.\ in contrast with the method of
    Mar{ê}ch{é} \emph{et al.}~\cite{mareche:2024-ispr}, whose
    convergence is not established. We propose here a normal estimator
    defined as the most orthogonal vector to the cone of visibility at
    the point of interest. Visibility is high in smooth regions and
    the induced visibility vectors are good approximations of tangent
    vectors. On the contrary the visibility is blocked at sharp
    features, so the normal vector is then not influenced by the
    geometry on the other side of the features. We demonstrate the
    superiority of this normal estimator for estimating curvatures
    along digital surfaces with sharp features.

@@

    \begin{figure}
        \begin{center}
            \begin{tabular}{|c||c|c|c|}
                \hline
                Curvature & II normals & CTriv normals & Our normals \\ \hline \hline
                \raisebox{18mm}{$\kappa_1$} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-k1-II.jpg} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-k1-CTriv.jpg}&
                \includegraphics[width=0.3\textwidth]{pictures/d20-k1-VN.jpg}\\ \hline
                \raisebox{18mm}{$\kappa_2$} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-k2-II.jpg} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-k2-CTriv.jpg}&
                \includegraphics[width=0.3\textwidth]{pictures/d20-k2-VN.jpg}\\ \hline
                \raisebox{18mm}{$H$} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-H-II.jpg} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-H-CTriv.jpg}&
                \includegraphics[width=0.3\textwidth]{pictures/d20-H-VN.jpg}\\ \hline
                \raisebox{18mm}{$G$} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-G-II.jpg} &
                \includegraphics[width=0.3\textwidth]{pictures/d20-G-CTriv.jpg}&
                \includegraphics[width=0.3\textwidth]{pictures/d20-G-VN.jpg}\\ \hline
            \end{tabular}
        \end{center}
        \caption{Estimation of curvatures using Corrected Normal
          Current estimators \cite{lachaud:2022-dcg} with a measure of
          radius $2$ for the shape ``Dice-20'' of Thingi10K database
          at resolution $256^3$. This estimator is parameterized by an
          input normal estimator: first column by ``Integral Invariant
          normals'' (radius $r=6$, second column by ``Convolved
          Trivial normals'' (radius $k=6$), third column by our
          presented normal estimator (deviation $\sigma=4$, radius
          $r=2\sigma=8$). Per row are displayed the estimated
          curvatures in order: first and second principal curvatures
          $\kappa_1$ and $\kappa_2$, mean curvature $H$, Gaussian
          curvature $G$. The range of curvature between deep blue and
          deep red is $\lbrack -0.1, 0.1 \rbrack$ with white color
          equal to $0$.}
    \end{figure}

>>>>

\usepackage{changes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{varwidth}
\usepackage{mathtools}
\usepackage{pgfplots}
\usepackage[dvipsnames]{xcolor}
\pgfplotsset{compat=1.18}
\usetikzlibrary {spy,arrows.meta,decorations.shapes}
\setauthormarkup{}
\definechangesauthor[name=jo, color=blue]{JO}
\definechangesauthor[name=rom, color=teal]{ROM}

@@

    Visibility is a fundamental concept in computational geometry and
    digital topology, with applications ranging from computer vision
    (occlusions), geometric modeling (feature detection,
    geodesics) to computer graphics (path tracing,
    shadowmap). Visibility within polygons in the plane has been
    extensively studied in the litterature~\cite{ghosh:2007-book} and
    is still studied in higher dimensional
    spaces~\cite{orourke:2017-book}. The general problem has a high
    complexity. For instance, the 3d visibility complex between $n$
    spheres and occluding spheres has a complexity
    $O(n^4)$~\cite{durand:2002-tog}. Exact 3d visibility between
    polygons and occluding polygons is often casted into 5d Plücker
    space~\cite{nirenstein:2002-ewr}, which nicely represents 3d lines
    and their respective positions. The computational cost remains
    high even with decomposition into coarse cells for speed up
    (several days of computation for 1M triangles at the time).

@@

    We focus on solving efficiently this problem because it is crucial
    in computing geometric normals along digital surfaces that are
    aware of salient features. Indeed, 2d discrete tangent estimators
    based on tangency or discrete line segments have proven to be both
    sensitive to sharp corners while multigrid convergent along the
    boundary of digitized smooth shapes~\cite{feschet:1999-dgci,lachaud:2007-ivc,nguyen:2011-pr}.
    Many ad hoc methods have been proposed for tangent plane or normal
    estimation on 3d digital surfaces
    (e.g. \cite{fourey:2009-cg,charrier:2011-iwcia,Cuel:2014-dgci,Lachaud:2017-lnm,mareche:2024-ispr}).
    If two of these methods~\cite{Cuel:2014-dgci,Lachaud:2017-lnm}
    have been shown to be multigrid convergent on digitization of
    shapes with smooth boundary, their formulation intrinsically
    smoothes features, e.g.\ in contrast with the method of
    Mar{ê}ch{é} \emph{et al.}~\cite{mareche:2024-ispr}, whose
    convergence is not established. We propose here a normal estimator
    defined as the most orthogonal vector to the cone of visibility at
    the point of interest. Visibility is high in smooth regions and
    the induced visibility vectors are good approximations of tangent
    vectors. On the contrary the visibility is blocked at sharp
    features, so the normal vector is not influenced by the
    geometry on the other side of the features. We demonstrate the
    superiority of this normal estimator for estimating curvatures
    along digital surfaces with sharp features.

@@

    \newcommand{\MyZoom}[1]{%
      \begin{tikzpicture}[spy using outlines={circle,magnification=1.8,size=2cm,connect spies}]
        \node[inner sep=0pt] {\pgfimage[width=0.3\textwidth]{#1}};
        \spy[overlay,blue] on (0.4,0.2) in node at (-0.8,0.8);
      \end{tikzpicture}}
    % \newcommand{\MyZoom}[1]{\includegraphics[width=0.3\textwidth]{#1}}

    \begin{figure}
        \begin{center}
            \begin{tabular}{|c||c|c|c|}
                \hline
                Curv. & II normals & CTriv normals & Our normals \\ \hline \hline
                \raisebox{18mm}{$\kappa_1$} &
                \MyZoom{pictures/d20-k1-II.jpg} &
                \MyZoom{pictures/d20-k1-CTriv.jpg}&
                \MyZoom{pictures/d20-k1-VN.jpg}\\ \hline
                \raisebox{18mm}{$\kappa_2$} &
                \MyZoom{pictures/d20-k2-II.jpg} &
                \MyZoom{pictures/d20-k2-CTriv.jpg}&
                \MyZoom{pictures/d20-k2-VN.jpg}\\ \hline
                \raisebox{18mm}{$H$} &
                \MyZoom{pictures/d20-H-II.jpg} &
                \MyZoom{pictures/d20-H-CTriv.jpg}&
                \MyZoom{pictures/d20-H-VN.jpg}\\ \hline
                \raisebox{18mm}{$G$} &
                \MyZoom{pictures/d20-G-II.jpg} &
                \MyZoom{pictures/d20-G-CTriv.jpg}&
                \MyZoom{pictures/d20-G-VN.jpg}\\ \hline
            \end{tabular}
        \end{center}
        \caption{Estimation of curvatures using Corrected Normal
          Current estimators \cite{lachaud:2022-dcg} with a measure of
          radius $2$ for the shape ``Dice-20'' of Thingi10K database
          at resolution $256^3$. This estimator is parameterized by an
          input normal estimator: first column by ``Integral Invariant
          normals'' (radius $r=6$, second column by ``Convolved
          Trivial normals'' (radius $k=6$), third column by our
          presented normal estimator (deviation $\sigma=4$, radius
          $r=2\sigma=8$). Per row are displayed the estimated
          curvatures in order: first and second principal curvatures
          $\kappa_1$ and $\kappa_2$, mean curvature $H$, Gaussian
          curvature $G$. The range of curvature between deep blue and
          deep red is $\lbrack -0.1, 0.1 \rbrack$ with white color
          equal to $0$.}
    \end{figure}
